---
title: "Reproducible Research Course Project 1"
author: "Deul Saha"
date: "26/10/2020"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(lubridate)
library(ggplot2)
```
## 1. Data Loading And Required Transformation.
```{r }
data<-read.csv("activity.csv")
data$date<-parse_date_time(data$date,"ymd")
```
## 2. Calculation for plotting of Histogram.
```{r fig.width=5, paged.print=TRUE}
df<-group_by(data,date)%>%summarise(steps=sum(steps,na.rm = TRUE))
g<-ggplot(df,aes(steps,fill="count"))
g+geom_histogram(bins  = 20)+ggtitle("Histogram for Total no of steps summed over days")+xlab("Steps")+theme_light()

```

## 3. Mean and median number of steps taken each day.
```{r paged.print=FALSE}
#Median 
md<-median(df$steps,na.rm = TRUE)
md
#Mean
mn<-mean(df$steps,na.rm = TRUE)
mn
```
## 4. Time series plot of the average number of steps taken.
```{r fig.width=6, message=TRUE, warning=TRUE, paged.print=TRUE}
tf<-group_by(data,interval)%>%summarise(steps=mean(steps,na.rm = TRUE))
tm<-ggplot(tf,aes(interval,steps))
tm+geom_line(lwd=0.5,color="black")+theme_light()+ggtitle(" No Of Steps Averaged Over Days For Different Intervals. ")+xlab("Interval")+ylab("Steps")
```

## 5. The 5-minute interval that, on average, contains the maximum number of steps.
```{r}
tf[which(tf$steps==max(tf$steps)),]
```
## 6. Code to describe and show a strategy for imputing missing data
The missing values for a particular interval is replaced by the mean of that interval.
```{r}
missing_activity<-data[is.na(data$steps),]
nonMissing_activity<-data[!is.na(data$steps),]
impu_activity<-left_join(missing_activity,tf,by="interval")%>%select(-steps.x)%>%rename(steps=steps.y)
cor_data<-rbind(nonMissing_activity,impu_activity)
```
## 7. Histogram of the total number of steps taken each day after missing values are imputed
```{r Histogram after Imputation, echo=TRUE, fig.width=5, message=TRUE, warning=TRUE, paged.print=TRUE}
df1<-group_by(cor_data,date)%>%summarise(steps=sum(steps))
 g_cor<-ggplot(df1,aes(steps,fill="count"))
 g_cor+geom_histogram(bins = 20)+ggtitle("Histogram of the total no of steps taken after data Imputation. ")+xlab("Steps")+ylab("count")
 
```

## Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends
```{r fig.height=4, fig.width=6, paged.print=TRUE}
#transformin the data 
z<-function(x){if(x==1|x==7){1} else {0}}
f<-sapply(wday(cor_data$date),z)
wd<-mutate(cor_data,weekdays=as.factor(f))
wd<-group_by(wd,interval,weekdays)%>% summarise(steps=mean(steps))
class(wd$weekdays)
levels(wd$weekdays)<-c("Weekdays","Weekend")
g<-ggplot(wd,aes(interval,steps))
g+geom_line()+facet_wrap(~weekdays)+ggtitle("Comaprision Among Weekdays And Weekend")+xlab("Interval")+ylab("Steps")
table(f)
#weekdays are coded as 0 and weekend as 1
```




